---
title: "Lab 6: Landscape Resistance, shared alleles"
author: "Norah Saarman"
date: "2024-10-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Section 13 Exercise

## a. Load libraries
```{r}
library(LandGenCourse)
#library(EcoGenetics)
library(GeNetIt)
library(hierfstat)
library(adegenet)
require(gstudio)       
require(dplyr)
require(tibble) 
require(sf)
require(popgraph)
require(RgoogleMaps)
require(geosphere)
require(proto)
require(sampling)
require(seqinr)
require(spacetime)
require(spdep)
require(here)
require(terra)
```

## d. Import rasters

```{r}
library(LandGenCourse)
library(GeNetIt)

RasterMaps <- terra::rast(system.file("extdata/covariates.tif", package="GeNetIt"))

plot(RasterMaps)
```
e. sites info (GPS coordinates for the study)
```{r}
data(ralu.site, package="GeNetIt")
sites <- ralu.site
str(sites)
```

2. Explore the data set... plot with terra::plot() function

a. Plot all rasters
```{r}
terra::plot(RasterMaps)
```

b. Plot spatial points over 'ffp' raster

```{r}
par(mar=c(2,2,1,1))
terra::plot(RasterMaps[["ffp"]])
terra::points(sites, pch=21, col="black", bg="white")
```
3. Setting cost values and calculating conductance

a. Resistance vs conductance values

b. scale rasters
```{r}
#cti <- terra::resample(cti, gsp, method= "bilinear")
```

c. Calculate conductance values
```{r}
RasterMaps[["err27"]]
```

```{r}
err.cost <- (1/RasterMaps[["err27"]])
err.cost
```
```{r}
RasterMaps[["ffp"]]
```

```{r}
ffp.cost <- (RasterMaps[["ffp"]]/5)
ffp.cost
```

```{r}
RasterMaps[["gsp"]]
```

```{r}
gsp.cost <- (RasterMaps[["gsp"]]-196)/15
gsp.cost
```

```{r}
RasterMaps[["cti"]]
```

```{r}
cti.cost <- RasterMaps[["cti"]]/5
cti.cost
```

d. Create a single landscape conductance raster

```{r}
cost1 <- (gsp.cost + cti.cost + err.cost + ffp.cost)
cost1
```
4. Convert conductance into effective distance
The higher the conductance, the lower the cost or resistance of a cell, and vice versa. We want to integrate conductance across cells to derive some measure of effective (or ecological) distance.

a. transition layer
```{r}
tr.cost1 <- gdistance::transition(raster::raster(cost1), transitionFunction=mean, directions=8) 
tr.cost1
```
b. visually inspect the raster

```{r}
par(mar=c(2,2,1,1))
raster::plot(raster::raster(tr.cost1))
```

c. Correct for geometric distortion

```{r}
tr.cost1 <- gdistance::geoCorrection(tr.cost1,type = "c",multpl=FALSE)
```

d. plot shortest paths
```{r}
sites.sp <- sf::as_Spatial(sites)

par(mar=c(2,2,1,2))
AtoB <- gdistance::shortestPath(tr.cost1, origin=sites.sp[1,], 
                                goal=sites.sp[2,], output="SpatialLines")
raster::plot(raster::raster(tr.cost1), xlab="x coordinate (m)", 
             ylab="y coordinate (m)",legend.lab="Conductance")
lines(AtoB, col="red", lwd=2)
points(sites.sp[1:2,])

par(mar=c(2,2,1,2))
raster::plot(raster::raster(tr.cost1), xlab="x coordinate (m)", 
             ylab="y coordinate (m)", legend.lab="Conductance")
points(sites.sp)

Neighbours <- spdep::tri2nb(sites.sp@coords, row.names = sites.sp$SiteName)

plot(Neighbours, sites.sp@coords, col="darkgrey", add=TRUE)
for(i in 1:length(Neighbours))
{
  for(j in Neighbours[[i]][Neighbours[[i]] > i])
  {
    AtoB <- gdistance::shortestPath(tr.cost1, origin=sites.sp[i,], 
                                goal=sites.sp[j,], output="SpatialLines")
    lines(AtoB, col="red", lwd=1.5)
  }
}
```
5. Create cost-distance matrices
a. Least cost distance

```{r}
cost1.dist <- gdistance::costDistance(tr.cost1,sites.sp)
```
b. Cost-distance matrix based on random paths (similar to Circuitscape)
```{r}
comm1.dist <- gdistance::commuteDistance(x = tr.cost1, coords = sites.sp)
```

c. Compare cost distances
```{r}
dist_df <- data.frame("cost1.dist"=as.numeric(cost1.dist),
                      "comm1.dist"=as.numeric(comm1.dist))

corr.LCD.comm <- cor(dist_df$cost1.dist, dist_df$comm1.dist, method = "spearman")
corr.LCD.comm

plot(cost1.dist~comm1.dist)
```
6. How does changing resolution affect these metrics?

a. Create a loop

```{r}
cor_cost <- c()
cor_comm <- c()
res_fact <- seq(2,20,2)
for(fac in res_fact){
  cost1_agg <- raster::aggregate(raster::raster(cost1), fact = fac)
  tr.cost_agg <- gdistance::transition(cost1_agg, 
                 transitionFunction=mean, directions=8)
  tr.cost_agg <- gdistance::geoCorrection(tr.cost_agg,type = "c",multpl=FALSE)
  cost.dist_agg <- gdistance::costDistance(tr.cost_agg, sites.sp)
  comm.dist_agg <- gdistance::commuteDistance(x = tr.cost_agg, coords = sites.sp)
  cost.dist_agg <- as.numeric(cost.dist_agg)
  comm.dist_agg <- as.numeric(comm.dist_agg)
  cor_cost <- c(cor_cost,cor(dist_df$cost1.dist, cost.dist_agg, 
                             method = "spearman"))
  cor_comm <- c(cor_comm,cor(dist_df$comm1.dist, comm.dist_agg, 
                             method = "spearman"))
}
```

b. Plot the results
```{r}
par(mar=c(4,4,1,1))
plot(y = cor_cost, x = res_fact, col = "red", pch = 19, 
     ylim = c(0.9,1), xlab = "Aggregation factor", ylab = "Spearman correlation")
points(y = cor_comm, x = res_fact, col = "blue", pch = 19)
legend("bottomleft", legend = c("Costdist","Commdist"), 
       pch = 19, col = c("red", "blue"))
```

