---
title: "Lab 2: Genetic Data"
author: "Norah Saarman"
date: "2024-09-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 4.3 Worked Example
https://bookdown.org/hhwagner1/LandGenCourse_book/WE_1.html

Make sure we have the packages, ONLY run once, and set to eval=FALSE.

```{r, eval = FALSE}
if(!require("adegenet")) install.packages("adegenet")
if(!requireNamespace("popgraph", quietly = TRUE))
{
  install.packages(c("RgoogleMaps", "geosphere", "proto", "sampling", 
                      "seqinr", "spacetime", "spdep"), dependencies=TRUE)
  remotes::install_github("dyerlab/popgraph")
}
if(!requireNamespace("gstudio", quietly = TRUE)) remotes::install_github("dyerlab/gstudio")

install.packages("here")
```

Load libraries (run every time) and hide text output when knit
```{r, results = FALSE, warning=FALSE, message=FALSE}
library(adegenet)
library(gstudio)
library(LandGenCourse)
library(tibble)
library(here)
library(vcfR)
library(pinfsc50)
library(utils)
```

## 1. Overview
The data set we will use is ralu.loci

Import straight from the package after library is loaded
```{r}
data("ralu.loci")
```

As a .csv file:

To download the csv file...
```{r,eval=FALSE}
if(!dir.exists(paste0(here(),"/downloads"))) dir.create(paste0(here(),"/downloads"))
file.copy(system.file("extdata", "ralu.loci.csv", package = "LandGenCourse"),
          paste0(here(), "/downloads/ralu.loci.csv"), overwrite=FALSE)
```
## 2. Import from csv file:

```{r}
Frogs <- read.csv(file="./downloads/ralu.loci.csv",header=TRUE)
as_tibble(Frogs)
```
Adding a column that gives us Frogs$FrogID
```{r}
Frogs <- data.frame(FrogID = paste(substr(Frogs$Pop, 1, 3), row.names(Frogs), sep="."), Frogs)
as_tibble(Frogs)

dim(Frogs)
```
## Some useful file directory functions
```{r}
#file.choose() but don't leave within code chunk without commenting it out

paste0(here(),"/output","/gobblygook")
```
## Save output file
```{r}
if(!dir.exists(paste0(here(),"/output"))) dir.create(paste0(here(),"/output"))

write.csv(ralu.loci, paste0(here(),"/output/ralu.loci.csv"), 
          quote=FALSE, row.names=FALSE)
```

## 3. Create a 'genind' object using adegenet
```{r}
Frogs[,c(4:11)]

Frogs.genind <- df2genind(X=Frogs[,c(4:11)], sep=":", ncode=NULL, ind.names= Frogs$FrogID, loc.names=NULL, pop=Frogs$Pop, NA.char="NA", ploidy=2, type="codom", strata=NULL, hierarchy=NULL)
```
## Check the genind object

```{r}
Frogs.genind
summary(Frogs.genind)
```
## 4. View info stored in 'genind' object

Pull out subsets of info:
```{r}
as_tibble(Frogs.genind@tab)
```
## 5. From here on, uses gstudio, and is OPTIONAL for now...

# Part 2: Section 4.4 R exercise Week 1

Do this section on your own!

Download file:
```{r}
file.copy(system.file("extdata", "pulsatilla_genotypes.csv", package = "LandGenCourse"),
          paste0(here(), "/downloads/pulsatilla_genotypes.csv"), overwrite=FALSE)
```

Import from csv as data frame:

```{r}
Flowers <- read.csv(paste0(here(), "/downloads/pulsatilla_genotypes.csv"), header=TRUE)
as_tibble(Flowers)
```

Create 'genind' object:
```{r}
Flowers.genind <- df2genind(X=Flowers[,c(6:19)], sep="\t", ncode=NULL, ind.names= Flowers$ID, loc.names=NULL, pop=Flowers$Population, NA.char="NA", ploidy=2, type="codom", strata=NULL, hierarchy=NULL)


```

Get info on genind object:
```{r}
Flowers.genind
```

Here, there is a problem because the genind object has 14 loci, but really this should be only 7 loci.

Going back to the Flowers dataframe and combining loci before creating the genind object
```{r}
as_tibble(Flowers)
```
I'll make a new data frame with the first 5 columns, then paste loc1_a:loc1_b, etc.
```{r}
Flowers <- data.frame(Flowers[,1:5],loc1 = paste(Flowers$loc1_a, Flowers$loc1_b, sep=":"), loc2 = paste(Flowers$loc2_a, Flowers$loc2_b, sep=":"), loc3 = paste(Flowers$loc3_a, Flowers$loc3_b, sep=":"), loc4 = paste(Flowers$loc4_a, Flowers$loc4_b, sep=":"), loc5 = paste(Flowers$loc5_a, Flowers$loc5_b, sep=":"), loc6 = paste(Flowers$loc6_a, Flowers$loc6_b, sep=":"), loc7 = paste(Flowers$loc7_a, Flowers$loc7_b, sep=":"))
as_tibble(Flowers)
```

Create 'genind' object:
```{r}
Flowers.genind <- df2genind(X=Flowers[,c(6:12)], sep=":", ncode=NULL, ind.names= Flowers$ID, loc.names=names(Flowers[,c(6:12)]), pop=Flowers$Population, NA.char="NA", ploidy=2, type="codom", strata=NULL, hierarchy=NULL)
```

Get info on genind object:
```{r}
Flowers.genind
summary(Flowers.genind)
```
Number of individuals per group/pop is 55-128.

## With Gstudio this time
```{r}
library(gstudio)
library(adegenet)

g.Flowers <- read_population("./downloads/pulsatilla_genotypes.csv",type = "column",locus.columns = c(6:19))

g.Flowers

g.Flowers.genind <- df2genind(X=g.Flowers[,c(6:12)], sep=":", ncode=NULL, ind.names=g.Flowers$ID, loc.names=NULL, pop=g.Flowers$Population, NA.char="", ploidy=2, type="codom", strata=NULL, hierarchy=NULL)

g.Flowers.genind

summary(g.Flowers.genind)
```
Compare the genind objects from both sources:
```{r}
summary(g.Flowers.genind)
summary(Flowers.genind)
```
