---
title: "Lab 4: Spatial Data, Tmap, Bubble Plots"
author: "Norah Saarman"
date: "2024-10-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Section 5.4 Exercise
https://bookdown.org/hhwagner1/LandGenCourse_book/r-exercise-week-2.html

## a. Load libraries
```{r, results = FALSE, warning=FALSE, message=FALSE}
require(adegenet)
require(LandGenCourse)
require(gstudio)       
require(dplyr)
require(tibble) 
require(sf)
```

## b. Import data with gstudio

First downloading the data
```{r,eval=FALSE}
if(!dir.exists(paste0(here(),"/downloads"))) dir.create(paste0(here(),"/downloads"))
file.copy(system.file("extdata", "ralu.loci.csv", package = "LandGenCourse"),
          paste0(here(), "/downloads/ralu.loci.csv"), overwrite=FALSE)
file.copy(system.file("extdata", "pulsatilla_genotypes.csv", package = "LandGenCourse"),
          paste0(here(), "/downloads/pulsatilla_genotypes.csv"), overwrite=FALSE)
```
Now loading the data with gstudio
```{r}
g.Flr <- read_population("./downloads/pulsatilla_genotypes.csv",type = "column",locus.columns = c(6:19))

#g.Flr
```

## c. Summarize by site

I would probably just use the function table(), but we can see where this takes us. 
```{r}
table(g.Flr$Population) 
```

Recommend using groups_by from library dplyr, with nested functions n and summarize, and to Write the result into a new object Pulsatilla

summarize(nIndiv = n(g.Flr ))

Example from Worked example:
pland_sum_b <- percentage_forest_500_df %>% 
  dplyr::group_by(plot_id) %>% 
  dplyr::summarize(sum_pland = sum(value))
pland_sum_b

```{r}
puls_df <- g.Flr %>%                 # create a df and piping
  dplyr::group_by(Population) %>%    # grouping by population and piping
  dplyr::summarize(nIndiv = n())     # summarize with n() and add name of column "nIndiv"
puls_df
```
## d. Add mean X and Y coordinates to object Pulsatilla

summarize(nIndiv = n(), myMean = n(myVar))
```{r}
puls_df <- g.Flr %>%                 # create a df and piping
  dplyr::group_by(Population) %>%    # grouping by population and piping
  dplyr::summarize(nIndiv = n(), meanX = mean(X), meanY = mean(Y))     # summarize with n() and add name of column 
puls_df
puls_df$meanX[2] == mean(g.Flr[g.Flr$Population == "A21",4])
```

## e. Convert to sf object

```{r}
Sites <- as.data.frame(puls_df[,2:4])
rownames(Sites) <- puls_df$Population
colnames(Sites) <- c("n","X","Y")
Pulsatilla <- st_as_sf(Sites, coords=c("X", "Y"))
```
## f. Specify the known projection 

The correct EPSG number for this dataset is: 31468. You can specify the CRS with: 
```{r}
st_crs(Pulsatilla) <- 31468

Pulsatilla
```

## g. Transform to lat/long projection
Adapt code from section 2.c to transform the projection to the “longlat” coordinate system, and write it into an object Pulsatilla.longlat.
```{r}
st_transform(Pulsatilla, crs = 4326)
Pulsatilla

plot(Pulsatilla)
```
## h. Create bubble plot
Adapt code from section 4.d to create a bubble plot of the number of individuals per population. Note: you may drop the argument key.entries as it has a default.

Here it is in live view mode, i don't need this in knitted output:
```{r, eval=FALSE}
library(tmap)
tmap_mode("view")
tm_shape(Pulsatilla) + tm_bubbles(col="n") 
```
Here it is again with plot mode so that I can have something in knitted output
```{r}
library(tmap)
tmap_mode("plot")
tm_shape(Pulsatilla) + tm_bubbles(col="n") 
```

## i. Save data as R object
Save the object Pulsatilla.longlat as an R object using the following code:
```{r,eval=FALSE}
saveRDS(Pulsatilla, file = "./output/Pulsatilla.longlat.rds")
```
We will need it for a later R exercise.


# BONUS material, working with 'sf' and 'terra'

##1. Load libraries, need to install GeNetIt
```{r}
library(LandGenCourse)
library(sf)
library(GeNetIt)
library(terra)
library(tmap)
library(dplyr)
library(tibble)
library(here)
```
##2. import/export ESRI shape files
a. Export ‘sf’ object to shapefile
```{r}
data(ralu.site)
if(!dir.exists(here("output"))) dir.create(here("output"))
dir.create(here("output/Sites"))
st_write(ralu.site, here("output/Sites/Sites.shp"), delete_dsn = TRUE)
```
b. Import shapefile to ‘sf’ object
```{r}
Sites.sf_a <- st_read(here("output/Sites/Sites.shp"))
Sites.sf_a 
plot(Sites.sf_a)
```
As sp
```{r}
data(ralu.site)
Sites.sp <- sf::as_Spatial(ralu.site)
Sites.sp
```
Convert back to sf
```{r}
Sites.sf_b <- sf::st_as_sf(Sites.sp)
Sites.sf_b
```

## b. Converting between terra and raster
```{r}
RasterMaps <- rast(system.file("extdata/covariates.tif", package="GeNetIt"))

RasterMaps.r <- raster::stack(RasterMaps)
RasterMaps.r 

plot(RasterMaps.r)
```
grab one layer
```{r}
gsp.r <- raster::raster(RasterMaps$gsp)
gsp.r 
plot(gsp.r)
```
Convert to "terra"
```{r}
gsp <- terra::rast(gsp.r)
gsp
plot(gsp)
plot(gsp.r)

RasterMaps_b <- terra::rast(RasterMaps.r)
RasterMaps_b
```

## 4. Plotting spatial data with tmap
a. Plotting just the geometry from the data stored in sf object
```{r}
data(ralu.site)
Sites.sf_c <- ralu.site
plot(st_geometry(Sites.sf_c))
```
b. Plot the attributes
```{r}
par(mar=c(2,2,2,2))
plot(Sites.sf_c, pch=16, cex=2)
```
```{r}

plot(Sites.sf_c[,c("Basin", "Depth_m")], pch=16)

plot(Sites.sf_c[,c(3,8)], pch=16)
```
c. bubble plot with tmap library
```{r}
tmap_mode("plot")
tm_shape(Sites.sf_c) + tm_bubbles(size="Depth_m", col="Basin") +
  tm_layout(legend.outside=TRUE, legend.outside.position="right") 
```
Extend extent of the plot
```{r}
Bbox = st_bbox(Sites.sf_c)
Bbox

#expand this by a factor of 0.1
delta.x <- Bbox[3] - Bbox[1]
delta.y <- Bbox[4] - Bbox[2]
Zoom <- 0.2
Bbox2 <- Bbox + c(-delta.x, -delta.y, delta.x, delta.y) * Zoom
Bbox2
```
```{r}
tmap_mode("plot")
Map1 <- tm_shape(Sites.sf_c, bbox=Bbox2) + 
  tm_bubbles(size="Depth_m", col="Basin") +
  tm_layout(legend.outside=TRUE, legend.outside.position="right") 
Map1
```
