---
title: "Lab 3: Genetic Diversity"
author: "Norah Saarman"
date: "2024-09-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 6.3 Worked Example
https://bookdown.org/hhwagner1/LandGenCourse_book/WE_3.html

Load libraries
```{r, results = FALSE, warning=FALSE, message=FALSE}
require(adegenet)
require(LandGenCourse)
require(pegas)       
require(PopGenReport)
require(dplyr)
require(poppr) 
require(here)
```

## 1. Overview
The data set we will use is ralu.loci


## 2. Import straight from the package after library is loaded
```{r}
data(ralu.loci, package="LandGenCourse")
Frogs <- data.frame(FrogID = paste(substr(ralu.loci$Pop, 1, 3), 
                                   row.names(ralu.loci), sep="."), ralu.loci)
Frogs.genind <- adegenet::df2genind(X=Frogs[,c(4:11)], sep=":", ncode=NULL, 
                          ind.names= Frogs$FrogID, loc.names=NULL, 
                          pop=Frogs$Pop, NA.char="NA", ploidy=2, 
                          type="codom", strata=NULL, hierarchy=NULL)
Frogs.genind
```

Get info on genind object, check that they are polymorphic
```{r}
Frogs.genind
summary(Frogs.genind)
```
Test for HWE with pegas:
```{r}
round(pegas::hw.test(Frogs.genind, B = 1000), digits = 3)
```

```{r}
# Chi-squared test: p-value
HWE.test <- data.frame(sapply(seppop(Frogs.genind), 
                              function(ls) pegas::hw.test(ls, B=0)[,3]))
HWE.test.chisq <- t(data.matrix(HWE.test))
{cat("Chi-squared test (p-values):", "\n")
round(HWE.test.chisq,3)}

# Monte Carlo: p-value
HWE.test <- data.frame(sapply(seppop(Frogs.genind), 
                              function(ls) pegas::hw.test(ls, B=1000)[,4]))
HWE.test.MC <- t(data.matrix(HWE.test))
{cat("MC permuation test (p-values):", "\n")
round(HWE.test.MC,3)}

alpha=0.05 # /96
Prop.loci.out.of.HWE <- data.frame(Chisq=apply(HWE.test.chisq<alpha, 2, mean), 
           MC=apply(HWE.test.MC<alpha, 2, mean))
Prop.loci.out.of.HWE             # Type this line again to see results table

Prop.pops.out.of.HWE <- data.frame(Chisq=apply(HWE.test.chisq<alpha, 1, mean), 
           MC=apply(HWE.test.MC<alpha, 1, mean))
Prop.pops.out.of.HWE             
```


```{r}
Chisq.fdr <- matrix(p.adjust(HWE.test.chisq,method="fdr"), 
                    nrow=nrow(HWE.test.chisq))
MC.fdr <- matrix(p.adjust(HWE.test.MC, method="fdr"), 
                    nrow=nrow(HWE.test.MC))

Prop.pops.out.of.HWE <- data.frame(Chisq=apply(HWE.test.chisq<alpha, 1, mean), 
           MC=apply(HWE.test.MC<alpha, 1, mean),
           Chisq.fdr=apply(Chisq.fdr<alpha, 1, mean),
           MC.fdr=apply(MC.fdr<alpha, 1, mean))
Prop.pops.out.of.HWE             
```

# Linkage Disequilibrium
```{r}
poppr::ia(Frogs.genind, sample=199)

LD.pair <- poppr::pair.ia(Frogs.genind)
LD.pair
```

# Null Alleles
```{r}
# Null alleles: depends on method! See help file.
Null.alleles <- PopGenReport::null.all(Frogs.genind)
```

# Genetic Diversity

## allelic richness

```{r}
Sum <- adegenet::summary(Frogs.genind)
names(Sum)
par(mar=c(5.5, 4.5,1,1))
barplot(Sum$pop.n.all, las=3, 
       xlab = "", ylab = "Number of alleles")

plot(Sum$n.by.pop, Sum$pop.n.all, 
       xlab = "Sample size", ylab = "Number of alleles")
abline(lm(Sum$pop.n.all ~ Sum$n.by.pop), col = "red")
```

```{r}
Richness <- PopGenReport::allel.rich(Frogs.genind, min.alleles = NULL)
Richness$alleles.sampled

par(mar=c(5.5, 4.5,1,1))
barplot(Richness$mean.richness, las=3, ylab="Rarefied allelic richness (Ar)")
plot(colMeans(Richness$pop.sizes), Richness$mean.richness,
     xlab="Valid sample size", 
     ylab="Rarefied allelic richness (Ar)")
abline(lm(Richness$mean.richness ~ colMeans(Richness$pop.sizes)), col="red")

```
No longer correlated with sample size.  
\pagebreak

# 6.4 Exercise

Task: Drop offspring (seeds, OffID==1) from dataset pulsatilla_genotypes.csv, check for HWE by site and locus and calculate Hexp for each site.

Practice indexing:
```{r}
vec <- 1:10
vec == 3 | vec == 6
vec[vec == 3 | vec == 6]
```


## Drop offspring
Using adegenet and base R
```{r}
library(adegenet)

# 1. CSV file "./downloads/pulsatilla_genotypes.csv" --> data frame 
# with base R function read.csv()
Flr <- read.csv("./downloads/pulsatilla_genotypes.csv", header=TRUE)

# 2. Select only adults with base R indexing of data frame 
# rows where OffID==0, all columns
Flr <- Flr[Flr$OffID==0,]

# 3. Combine columns with base R function paste()
Flr <- data.frame(Flr[,1:5],loc1 = paste(Flr[,6],  Flr[,7], sep=":"), 
                            loc2 = paste(Flr[,8],  Flr[,9], sep=":"), 
                            loc3 = paste(Flr[,10], Flr[,11], sep=":"), 
                            loc4 = paste(Flr[,12], Flr[,13], sep=":"), 
                            loc5 = paste(Flr[,14], Flr[,15], sep=":"),  
                            loc6 = paste(Flr[,16], Flr[,17], sep=":"),  
                            loc7 = paste(Flr[,18], Flr[,19], sep=":"))
# 4. Create genind object with "adegenet" function df2genind() 
# using NA.char = "NA"
Flr.genind <- df2genind(X=Flr[,c(6:12)], sep=":", ncode=NULL,
                        ind.names= Flr$ID, loc.names=names(Flr[,c(6:12)]), pop=Flr$Population,
                        NA.char="NA", ploidy=2, type="codom", strata=NULL, hierarchy=NULL)

# 5. Check genind object
Flr.genind
summary(Flr.genind)
```

# Count the number of individuals in each pop
```{r}
table(Flr$Population)
```

## With Gstudio this time
```{r}
library(gstudio)
library(adegenet)

# 1. CSV file "./downloads/pulsatilla_genotypes.csv" --> data frame 
# with "gstudio" function read_population()
g.Flr <- read_population("./downloads/pulsatilla_genotypes.csv",
                       type = "column",locus.columns = c(6:19))

# 2. Select only adults with base R indexing of data frame 
# rows where OffID==0, all columns
g.Flr <- g.Flr[g.Flr$OffID==0,]

# 3. Nothing to do here

# 4. Create genind object with "adegenet" function df2genind() 
# using NA.char = ""
g.Flr.genind <- df2genind(X=g.Flr[,c(6:12)], sep=":", ncode=NULL, ind.names=g.Flr$ID, loc.names=NULL, pop=g.Flr$Population, NA.char="", ploidy=2, type="codom", strata=NULL, hierarchy=NULL)

# 5. Check genind object
g.Flr.genind
summary(g.Flr.genind)
```
## Check for HWE by site and locus and calculate Hexp for each site

Test for HWE with pegas by site
```{r}
# Chi-squared test: p-value
HWE.test <- data.frame(sapply(seppop(g.Flr.genind), 
                              function(ls) pegas::hw.test(ls, B=0)[,3]))
HWE.test.chisq <- t(data.matrix(HWE.test))
{cat("Chi-squared test (p-values):", "\n")
round(HWE.test.chisq,3)}

# Monte Carlo: p-value
HWE.test <- data.frame(sapply(seppop(g.Flr.genind), 
                              function(ls) pegas::hw.test(ls, B=1000)[,4]))
HWE.test.MC <- t(data.matrix(HWE.test))
{cat("MC permuation test (p-values):", "\n")
round(HWE.test.MC,3)}

alpha=0.05 # /96
Prop.loci.out.of.HWE <- data.frame(Chisq=apply(HWE.test.chisq<alpha, 2, mean), 
           MC=apply(HWE.test.MC<alpha, 2, mean))
Prop.loci.out.of.HWE             # Type this line again to see results table

Prop.pops.out.of.HWE <- data.frame(Chisq=apply(HWE.test.chisq<alpha, 1, mean), 
           MC=apply(HWE.test.MC<alpha, 1, mean))
Prop.pops.out.of.HWE             
```


```{r}
Chisq.fdr <- matrix(p.adjust(HWE.test.chisq,method="fdr"), 
                    nrow=nrow(HWE.test.chisq))
MC.fdr <- matrix(p.adjust(HWE.test.MC, method="fdr"), 
                    nrow=nrow(HWE.test.MC))

Prop.pops.out.of.HWE <- data.frame(Chisq=apply(HWE.test.chisq<alpha, 1, mean), 
           MC=apply(HWE.test.MC<alpha, 1, mean),
           Chisq.fdr=apply(Chisq.fdr<alpha, 1, mean),
           MC.fdr=apply(MC.fdr<alpha, 1, mean))
Prop.pops.out.of.HWE   
```
# Hexp for each site

```{r}
Hobs <- t(sapply(seppop(g.Flr.genind), function(ls) summary(ls)$Hobs))
Hexp <- t(sapply(seppop(g.Flr.genind), function(ls) summary(ls)$Hexp))
{cat("Expected heterozygosity (Hexp):", "\n")
round(Hexp, 2)}
```
```{r}
  par(mar=c(5.5, 4.5, 1, 1))
  Hobs.pop <- apply(Hobs, MARGIN = 1, FUN = mean)
  Hexp.pop <- apply(Hexp, 1, mean) 
  barplot(Hexp.pop, ylim=c(0,1), las=3, ylab="Avg. Expected heterozygosity", xlab = "Site")
  barplot(Hobs.pop, ylim=c(0,1), las=3, ylab="Avg. Observed heterozygosity", xlab = "Site")
```

Data frame of summary statistics
```{r}
Sum <- summary(g.Flr.genind)
H.pop <- data.frame(Pop = names(Hobs.pop),
                    n = Sum[2]$n.by.pop,
                              Hobs = Hobs.pop,
                              Hexp = Hexp.pop)
H.pop
```

Save "H.pop" for later in output file (eval = FALSE so that it doesn't do this every time I knit) 
```{r, eval = FALSE}
saveRDS(H.pop, file = "./output/H.pop.rds")
```
## Question: Which site had the lowest expected heterozygosity? 

Site G05a had the lowest average expected heterozygosity, at 0.6503175.

