---
title: "Lab 5: Spatial Statistics"
author: "Norah Saarman"
date: "2024-10-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Section 5.4 Exercise
https://bookdown.org/hhwagner1/LandGenCourse_book/r-exercise-week-2.html

## a. Load libraries
```{r}
library(LandGenCourse)
#library(EcoGenetics)
library(GeNetIt)
library(hierfstat)
library(adegenet)
require(gstudio)       
require(dplyr)
require(tibble) 
require(sf)
require(popgraph)
require(RgoogleMaps)
require(geosphere)
require(proto)
require(sampling)
require(seqinr)
require(spacetime)
require(spdep)
require(here)
```

## b. Import to adegenet object G-Studio, Drop offspring:
```{r}
# 1. CSV file "./downloads/pulsatilla_genotypes.csv" --> data frame 
# with "gstudio" function read_population()
Flr <- read_population("./downloads/pulsatilla_genotypes.csv",
                       type = "column",locus.columns = c(6:19))

# 2. Select only adults with base R indexing of data frame 
# rows where OffID==0, all columns
Flr <- Flr[Flr$OffID==0,]

# 3. Nothing to do here

# 4. Create genind object with "adegenet" function df2genind() 
# using NA.char = ""
Flr.genind <- df2genind(X=Flr[,c(6:12)], sep=":", ncode=NULL, ind.names=Flr$ID, loc.names=NULL, pop=Flr$Population, NA.char="", ploidy=2, type="codom", strata=NULL, hierarchy=NULL)

# 5. Check genind object
Flr.genind
summary(Flr.genind)
```

## c. Plot locations of individuals from site A25 From data frame
```{r}
# Select data for Population "A25"
Sites <- Flr[Flr$Population == "A25", c("X", "Y")]

# Check if the data frame has valid coordinates
head(Sites)

# Convert to an 'sf' object
Sites_sf <- st_as_sf(Sites, coords = c("X", "Y"), crs = 31468) 
# Transform to WGS84 (lat/long, EPSG:4326)
Sites_latlon <- st_transform(Sites_sf, crs = 4326)

plot(Sites_latlon)

```  
  
## d. Add geographic info to genind object

Genind object @other can hold a list such as spatial coordinates
```{r}
# Select xy coordinates from Flr
Sites <- Flr[, c("X", "Y")]

# Convert to an 'sf' object with the correct original CRS (replace 31468 if needed)
Sites_sf <- st_as_sf(Sites, coords = c("X", "Y"), crs = 31468)

# Transform to WGS84 (lat/lon, EPSG:4326)
Sites_latlon <- st_transform(Sites_sf, crs = 4326)

# Extract coordinates from the transformed sf object
latlon_matrix <- st_coordinates(Sites_latlon)

# Convert to a matrix with column names "longitude" and "latitude"
colnames(latlon_matrix) <- c("longitude", "latitude")

# Add the matrix to the @other slot with the name "xy"
Flr.genind@other <- list(xy = latlon_matrix)
```

## e. Calculate genetic and geographic euclidean distance (for Site A25)
I can't tell if we are meant to proceed with just population A25? If so, I would first subset like this:
```{r}
# Subset only site A25
Flr.A25 <- Flr.genind[Flr.genind@pop == "A25"]
```
 
For now I'll proceed with the full genind object.

### Step 1: Genetic distance with adegenet::propShared:
```{r}
# Calculate genetic distance matrix
Dgen_matrix <- adegenet::propShared(Flr.A25)

# Convert to a distance vector (1 - proportion shared)
Dgen <- as.dist(1 - Dgen_matrix)
Dgen.vec <- as.vector(Dgen)
```

### Step 2: Geographic with dist()

### One option is to use the @other slot for lat long coordinates
```{r}
# Extract the 'xy' matrix from the @other slot
xy_matrix <- Flr.A25@other$xy

# Convert the matrix back into a data frame
df_xy <- as.data.frame(xy_matrix)
```

### Other (simpler) option is to pull from the original data frame, this is in the grid coordinate system
```{r}
df_xy <- Flr[Flr$Population == "A25", c("X","Y")]
```

```{r}
colnames(df_xy) <- c("X", "Y") 

# Calculate Euclidean geographic distances
Dgeo <- dist(df_xy)

Dgeo.vec <- as.vector(Dgeo)
```

### Adapt section 4a to Visualize
```{r}
# Load necessary packages
library(MASS)  # For kde2d()
library(scales)  # For transparent colors using alpha

# Step 5: Check visual linearity
par(mar = c(4, 4, 0, 0))  # Adjust plot margins

# Generate density for better visualization
dens <- kde2d(Dgeo.vec, Dgen.vec, n = 300)

# CrDgeo# Create a color palette
myPal <- colorRampPalette(c("white", "blue", "gold", "orange", "red"))

# Plot geographic vs genetic distance
plot(Dgeo.vec, Dgen.vec, pch = 20, cex = 0.5,  
     xlab = "Geographic Distance", ylab = "Genetic Distance")

# Add density image with transparency
image(dens, col = alpha(myPal(300), 0.7), add = TRUE)

# Add linear regression line
abline(lm(Dgen.vec ~ Dgeo.vec), col = "black")

# Add loess smoothing line in red
lines(loess.smooth(Dgeo.vec, Dgen.vec), col = "red")
```
Do you notice something unusual in the plot? Why are there so few different values of genetic distance?

We have 56 individuals, and 7 loci, 105 alleles... Maybe low heterozygosity? No. Maybe closely related or even clonal individuals because of vegetative reproduction?

Do you think there is spatial autocorrelation? If so, up to what distance? Yes, it looks like there is correlation of Dgeo and Dgen up until ~ 15 geographic units.

## f. Mantel test 
```{r}

IBD <- vegan::mantel(Dgen,Dgeo, method="pearson")
IBD
```
## g. Mantel Correlogram Without EcoGenetics

### Step 3: Define Distance Classes (Binning Geographic Distances)
To create a Mantel correlogram, we need to divide the geographic distances into bins or classes.

```{r}
# Step 3: Create geographic distance bins (classes)
# Define the number of bins (adjust this as necessary)
num_bins <- 10

# Get the range of geographic distances
max_dist <- max(Dgeo.vec, na.rm = TRUE)

# Define the breaks for binning distances
breaks <- seq(0, max_dist, length.out = num_bins + 1)

# Assign each pairwise distance to a bin (using cut)
Dgeo_bins <- cut(as.vector(Dgeo), breaks = breaks, include.lowest = TRUE)
```

### Step 4. Perform Mantel Tests for Each Distance Class
In each bin, we perform a Mantel test to assess the correlation between genetic and geographic distances.

```{r}
# Step 4: Perform Mantel tests for each distance bin
library(vegan)

# Initialize a vector to store the Mantel correlation coefficients
mantel_results <- numeric(num_bins)

# Loop over each bin and perform a Mantel test
for (i in 1:num_bins) {
  # Create a mask for the current bin (TRUE for distances in this bin)
  bin_mask <- as.matrix(Dgeo) >= breaks[i] & as.matrix(Dgeo) < breaks[i + 1]

  # Skip bins with no data
  if (sum(bin_mask) == 0) next

  # Perform Mantel test for this bin
  mantel_test <- mantel(as.dist(bin_mask), as.dist(Dgen), 
                        method = "pearson", permutations = 199)

  # Store the Mantel statistic (correlation)
  mantel_results[i] <- mantel_test$statistic
}
```

### Step 5: Plot the Mantel Correlogram
We plot the Mantel correlation coefficients across the distance bins to visualize the relationship.

```{r}
# Step 5: Plot the Mantel correlogram
plot(1:num_bins, mantel_results, type = "b", pch = 16, col = "blue",
     xlab = "Distance Class", ylab = "Mantel Correlation",
     main = "Mantel Correlogram: Genetic vs Geographic Distance")

# Add a horizontal line at 0 for reference
abline(h = 0, lty = 2, col = "red")
```


**Questions:** What is the range of spatial autocorrelation in P. vulgaris in site A25?

Based on a plot of genetic distance against Euclidean distance?

Based on where the Mantel correlogram reaches 0?

Based on statistical significance tests for the Mantel correlogram (with default settings: one-sided alternative “less”, Holm’s adjustment)?

# Interpretation:

Range of Spatial Autocorrelation from the Mantel Correlogram
Mantel Correlogram Interpretation:
The Mantel correlogram shows Mantel correlation coefficients plotted across geographic distance bins. This allows you to assess at which spatial scales genetic similarity (spatial autocorrelation) is strongest.

Range of Spatial Autocorrelation:

The range is the largest geographic distance bin where the Mantel correlation coefficient is significantly greater than 0 (positive autocorrelation).
When the Mantel coefficient reaches 0 (or dips below it), this marks the end of spatial autocorrelation, meaning that at larger distances, genetic similarity is no longer influenced by geographic proximity.

Using Statistical Significance to Identify the Range
One-Sided Test (Alternative = "less"):
The one-sided alternative hypothesis (alternative = "less") tests whether there is positive spatial autocorrelation, i.e., whether genetic distance increases with geographic distance. This is relevant for isolation-by-distance (IBD) models.

Holm’s Adjustment for Multiple Tests:
Since the Mantel correlogram involves multiple distance classes, Holm’s adjustment ensures that statistical significance is corrected for multiple comparisons, reducing the risk of false positives.

How to Interpret Statistical Significance:

For each distance bin, check if the p-value (adjusted using Holm’s method) is statistically significant (e.g., p < 0.05).
Significant positive Mantel correlation (with alternative = "less") in early bins indicates that genetic similarity is highest at smaller geographic scales.
The range of spatial autocorrelation is the largest distance bin where the Mantel test remains significant. Beyond this range, genetic similarity no longer follows a geographic pattern.
