---
title: "Lab 4: Spatial Data"
author: "Norah Saarman"
date: "2024-10-03"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Section 5.4 Exercise
https://bookdown.org/hhwagner1/LandGenCourse_book/r-exercise-week-2.html

## a. Load libraries
```{r, results = FALSE, warning=FALSE, message=FALSE}
require(adegenet)
require(LandGenCourse)
require(gstudio)       
require(dplyr)
require(tibble) 
require(sf)
```

## b. Import data with gstudio

First downloading the data
```{r,eval=FALSE}
if(!dir.exists(paste0(here(),"/downloads"))) dir.create(paste0(here(),"/downloads"))
file.copy(system.file("extdata", "ralu.loci.csv", package = "LandGenCourse"),
          paste0(here(), "/downloads/ralu.loci.csv"), overwrite=FALSE)
file.copy(system.file("extdata", "pulsatilla_genotypes.csv", package = "LandGenCourse"),
          paste0(here(), "/downloads/pulsatilla_genotypes.csv"), overwrite=FALSE)
```
Now loading the data with gstudio
```{r}
g.Flr <- read_population("./downloads/pulsatilla_genotypes.csv",type = "column",locus.columns = c(6:19))

g.Flr
```

## c. Summarize by site

I would probably just use the function table(), but we can see where this takes us. 
```{r}
table(g.Flr$Population) 
```

Recommend using groups_by from library dplyr, with nested functions n and summarize, and to Write the result into a new object Pulsatilla

summarize(nIndiv = n(g.Flr ))

Example from Worked example:
pland_sum_b <- percentage_forest_500_df %>% 
  dplyr::group_by(plot_id) %>% 
  dplyr::summarize(sum_pland = sum(value))
pland_sum_b

```{r}
puls_df <- g.Flr %>%                 # create a df and piping
  dplyr::group_by(Population) %>%    # grouping by population and piping
  dplyr::summarize(nIndiv = n())     # summarize with n() and add name of column "nIndiv"
puls_df
```
## d. Add mean X and Y coordinates to object Pulsatilla

summarize(nIndiv = n(), myMean = n(myVar))
```{r}
puls_df <- g.Flr %>%                 # create a df and piping
  dplyr::group_by(Population) %>%    # grouping by population and piping
  dplyr::summarize(nIndiv = n(), meanX = mean(X), meanY = mean(Y))     # summarize with n() and add name of column 
puls_df
puls_df$meanX[2] == mean(g.Flr[g.Flr$Population == "A21",4])
```

## e. Convert to sf object

```{r}
Sites <- as.data.frame(puls_df[,2:4])
rownames(Sites) <- puls_df$Population
colnames(Sites) <- c("n","X","Y")
Pulsatilla <- st_as_sf(Sites, coords=c("X", "Y"))
```
## f. Specify the known projection 

The correct EPSG number for this dataset is: 31468. You can specify the CRS with: 
```{r}
st_crs(Pulsatilla) <- 31468

Pulsatilla
```

## g. Transform to lat/long projection
Adapt code from section 2.c to transform the projection to the “longlat” coordinate system, and write it into an object Pulsatilla.longlat.
```{r}
st_transform(Pulsatilla, crs = 4326)
Pulsatilla

plot(Pulsatilla)
```
## h. Create bubble plot
Adapt code from section 4.d to create a bubble plot of the number of individuals per population. Note: you may drop the argument key.entries as it has a default.
```{r}
tmap_mode("view")
tm_shape(Pulsatilla) + tm_bubbles(col="n") 
```

## i. Save data as R object
Save the object Pulsatilla.longlat as an R object using the following code:
saveRDS(Pulsatilla.longlat, file = here::here("output/Pulsatilla.longlat.rds")).
We will need it for a later R exercise.
