---
title: "Lab 2"
author: "Norah Saarman"
date: "2024-09-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 4.3 Worked Example
https://bookdown.org/hhwagner1/LandGenCourse_book/WE_1.html

The data set we will use is ralu.loci

Make sure we have the packages:

if(!require("adegenet")) install.packages("adegenet")
if(!requireNamespace("popgraph", quietly = TRUE))
{
  install.packages(c("RgoogleMaps", "geosphere", "proto", "sampling", 
                      "seqinr", "spacetime", "spdep"), dependencies=TRUE)
  remotes::install_github("dyerlab/popgraph")
}
if(!requireNamespace("gstudio", quietly = TRUE)) remotes::install_github("dyerlab/gstudio")

install.packages("here")

Load libraries

```{r}
library(adegenet)
library(gstudio)
library(LandGenCourse)
library(tibble)
library(here)
library(vcfR)
library(pinfsc50)
library(utils)
```
## Import data

straight from the package after library is loaded
```{r}
data("ralu.loci")
```

As a .csv file:

To download the csv file...
```{r}
if(!dir.exists(paste0(here(),"/downloads"))) dir.create(paste0(here(),"/downloads"))
file.copy(system.file("extdata", "ralu.loci.csv", package = "LandGenCourse"),
          paste0(here(), "/downloads/ralu.loci.csv"), overwrite=FALSE)
```
Import from csv file:

```{r}
Frogs <- read.csv(file="./downloads/ralu.loci.csv",header=TRUE)
as_tibble(Frogs)
```
Adding a column that gives us Frogs$FrogID
```{r}
Frogs <- data.frame(FrogID = paste(substr(Frogs$Pop, 1, 3), row.names(Frogs), sep="."), Frogs)
as_tibble(Frogs)

dim(Frogs)
```
## Some useful file directory functions
```{r}
here()
#file.choose() but don't leave within code chunk without commenting it out

paste0(here(),"/output","/gobblygook")
```
## save output file

```{r}
if(!dir.exists(paste0(here(),"/output"))) dir.create(paste0(here(),"/output"))

write.csv(ralu.loci, paste0(here(),"/output/ralu.loci.csv"), 
          quote=FALSE, row.names=FALSE)
```

## Create a 'genind' object using adegenet
```{r}
Frogs[,c(4:11)]

Frogs.genind <- df2genind(X=Frogs[,c(4:11)], sep=":", ncode=NULL, ind.names= Frogs$FrogID, loc.names=NULL, pop=Frogs$Pop, NA.char="NA", ploidy=2, type="codom", strata=NULL, hierarchy=NULL)

```
## Check the genind object

```{r}
Frogs.genind
summary(Frogs.genind)
```
## Pull out subsets of info:
```{r}
as_tibble(Frogs.genind@tab)
```
# From 5 on uses gstudio, and is optional for now.

## Section 4.4 R exercise Week 1
