---
title: "Lab 3"
author: "Norah Saarman"
date: "2024-09-17"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 6.3 Worked Example
https://bookdown.org/hhwagner1/LandGenCourse_book/WE_3.html

Load libraries
```{r}
#require(adegenet)
require(LandGenCourse)
#require(pegas)       
require(PopGenReport)
require(dplyr)
require(poppr) 
```

## 1. Overview
The data set we will use is ralu.loci


## 2. Import straight from the package after library is loaded
```{r}
data(ralu.loci, package="LandGenCourse")
Frogs <- data.frame(FrogID = paste(substr(ralu.loci$Pop, 1, 3), 
                                   row.names(ralu.loci), sep="."), ralu.loci)
Frogs.genind <- adegenet::df2genind(X=Frogs[,c(4:11)], sep=":", ncode=NULL, 
                          ind.names= Frogs$FrogID, loc.names=NULL, 
                          pop=Frogs$Pop, NA.char="NA", ploidy=2, 
                          type="codom", strata=NULL, hierarchy=NULL)
Frogs.genind
```

Get info on genind object, check that they are polymorphic
```{r}
Frogs.genind
summary(Frogs.genind)
```
Test for HWE with pegas:
```{r}
round(pegas::hw.test(Frogs.genind, B = 1000), digits = 3)
```

```{r}
# Chi-squared test: p-value
HWE.test <- data.frame(sapply(seppop(Frogs.genind), 
                              function(ls) pegas::hw.test(ls, B=0)[,3]))
HWE.test.chisq <- t(data.matrix(HWE.test))
{cat("Chi-squared test (p-values):", "\n")
round(HWE.test.chisq,3)}

# Monte Carlo: p-value
HWE.test <- data.frame(sapply(seppop(Frogs.genind), 
                              function(ls) pegas::hw.test(ls, B=1000)[,4]))
HWE.test.MC <- t(data.matrix(HWE.test))
{cat("MC permuation test (p-values):", "\n")
round(HWE.test.MC,3)}

alpha=0.05 # /96
Prop.loci.out.of.HWE <- data.frame(Chisq=apply(HWE.test.chisq<alpha, 2, mean), 
           MC=apply(HWE.test.MC<alpha, 2, mean))
Prop.loci.out.of.HWE             # Type this line again to see results table

Prop.pops.out.of.HWE <- data.frame(Chisq=apply(HWE.test.chisq<alpha, 1, mean), 
           MC=apply(HWE.test.MC<alpha, 1, mean))
Prop.pops.out.of.HWE             
```


```{r}
Chisq.fdr <- matrix(p.adjust(HWE.test.chisq,method="fdr"), 
                    nrow=nrow(HWE.test.chisq))
MC.fdr <- matrix(p.adjust(HWE.test.MC, method="fdr"), 
                    nrow=nrow(HWE.test.MC))

Prop.pops.out.of.HWE <- data.frame(Chisq=apply(HWE.test.chisq<alpha, 1, mean), 
           MC=apply(HWE.test.MC<alpha, 1, mean),
           Chisq.fdr=apply(Chisq.fdr<alpha, 1, mean),
           MC.fdr=apply(MC.fdr<alpha, 1, mean))
Prop.pops.out.of.HWE             
```

# Linkage Disequilibrium
```{r}
poppr::ia(Frogs.genind, sample=199)

LD.pair <- poppr::pair.ia(Frogs.genind)
LD.pair
```

# Null Alleles
```{r}
# Null alleles: depends on method! See help file.
Null.alleles <- PopGenReport::null.all(Frogs.genind)
```

# Genetic Diversity

## allelic richness

```{r}
Sum <- adegenet::summary(Frogs.genind)
names(Sum)
par(mar=c(5.5, 4.5,1,1))
barplot(Sum$pop.n.all, las=3, 
       xlab = "", ylab = "Number of alleles")

plot(Sum$n.by.pop, Sum$pop.n.all, 
       xlab = "Sample size", ylab = "Number of alleles")
abline(lm(Sum$pop.n.all ~ Sum$n.by.pop), col = "red")
```

```{r}
Richness <- PopGenReport::allel.rich(Frogs.genind, min.alleles = NULL)
Richness$alleles.sampled

par(mar=c(5.5, 4.5,1,1))
barplot(Richness$mean.richness, las=3, ylab="Rarefied allelic richness (Ar)")
plot(colMeans(Richness$pop.sizes), Richness$mean.richness,
     xlab="Valid sample size", 
     ylab="Rarefied allelic richness (Ar)")
abline(lm(Richness$mean.richness ~ colMeans(Richness$pop.sizes)), col="red")

```

# 6.4 Exercise

```{r}
Flowers <- read.csv(paste0(here(), "/downloads/pulsatilla_genotypes.csv"), header=TRUE)
as_tibble(Flowers)
```

## Filter offspring (seeds) from the dataset
```{r}
Flowers <- filter(Flowers, OffID==0)
as_tibble(Flowers)
```

## Count the number of individuals in each pop
Adults.by.site<-split(pulsatilla_adults, pulsatilla_adults$Population, T)
sapply(Adults.by.site, count)

Dataframe with the first 5 columns, then paste loc1_a:loc1_b, etc.
```{r}
Flowers <- data.frame(Flowers[,1:5],loc1 = paste(Flowers$loc1_a, Flowers$loc1_b, sep=":"), loc2 = paste(Flowers$loc2_a, Flowers$loc2_b, sep=":"), loc3 = paste(Flowers$loc3_a, Flowers$loc3_b, sep=":"), loc4 = paste(Flowers$loc4_a, Flowers$loc4_b, sep=":"), loc5 = paste(Flowers$loc5_a, Flowers$loc5_b, sep=":"), loc6 = paste(Flowers$loc6_a, Flowers$loc6_b, sep=":"), loc7 = paste(Flowers$loc7_a, Flowers$loc7_b, sep=":"))
as_tibble(Flowers)
```

Create 'genind' object:
```{r}
Flowers.genind <- df2genind(X=Flowers[,c(6:12)], sep=":", ncode=NULL, ind.names= Flowers$ID, loc.names=names(Flowers[,c(6:12)]), pop=Flowers$Population, NA.char="NA", ploidy=2, type="codom", strata=NULL, hierarchy=NULL)
```

Get info on genind object:
```{r}
Flowers.genind
summary(Flowers.genind)
```

